Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2019-05-07 icdm/submission/includes/discussion.tex
summary and transition evaluation
discussion improved

\section{Discussion}\label{chap:discussion}

%Discuss the results. What is the outcome of your experimetns?
This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some differences. It is followed by a subsection describing the limitations and possible extensions of this work. This subsection is concluded by the problems we encountered during the implementation so others can avoid them in the future.

\subsection{Our work}
This subsection discusses our findings, first we will analyze how input shuffling could influence the learning rate and conclude with the discussion of unexpected results.

\subsubsection{Shuffling and learning rate with SGD}
When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag} Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers. One possibility is the use of the batched version of the SGNS. In consequence, when the input is not shuffled the same word will often appear in one batch, hence the optimizer is presented with an average value of the gradient, which can be imprecise. This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate.

@@

\subsection{Comparison to Gensim}
This subsection will compare our finding extensively to Gensim \cite{gensim}. As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations. Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.

@@

\subsubsection{Configuration of Gensim}
The training with Gensim has a lot of possible parameters an extended list can be found in the appendix. This subsection will only describe the parameters we changed from the default setting. The description of each parameter has the following form: \\
\texttt{name} (type) -- \textit{Description} -- Value\\
Parameters:
\begin{itemize}
\item \texttt{sentences} (iterable of iterables) -- \textit{Dataset} -- text8 document splitted into sentences of 20 words
\item \texttt{size}(int) – \textit{Dimensionality of the word vectors } -- 100
\item \texttt{window} (int) --\textit {Maximum distance between the current and predicted word within a sentence }-- 5
\item \texttt{min\_count }(int) --\textit{ Ignores all words with total frequency lower than this }-- 5
\item \texttt{workers} (int) -- \textit{ Number of threads used to train the model} -- 4
\item \texttt{sg} ({0, 1}) --\textit{ Training algorithm: 1 for skip--gram}--1
\item \texttt{negative} (int) --\textit{Number of negative samples}-- 10
\item \texttt{ ns\_exponent} (float) --\textit{ Exponent in the unigram distribution, when choosing random samples, as shown in Equation \ref{eq:unigram} }-- 0.75
\item \texttt{alpha} (float) --\textit{ The initial learning rate. }-- 0.025
\item \texttt{min\_alpha} (float) --\textit{ Learning rate will linearly drop to min\_alpha as training progresses }--0.0001
\item \texttt{sample} (float) --\textit{ Treshold for subsampling as described in Equation \ref{eq:sampling}} -- 1e--4
\item \texttt{iter} (int) --\textit{ Number of iterations (epochs) over the corpus. }-- 10
\
\item \texttt{compute\_loss} (bool) --\textit{If True, loss is stored at the end of each batch}-- True
\item \texttt{callbacks} (iterable of CallbackAny2Vec) --\textit{ Set of functions executed in order to follow the loss and the progress of the model in word similarity }-- see Appendix
\end{itemize}

@@

\subsection{Future Work}
This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers. As with every work, there still exists possible extensions. First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset. The fact that we only used one dataset as well that it's a small one is problematic. By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words. There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field. But the main issue with using only one data set it that it has been shown that some optimizers perform better with specific shapes of loss functions. To make a compelling argument it's, therefore, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently, it needs to be confirmed with other datasets as well.
Furthermore, our implementation did not outperform Gensim in run time, as this was not the goal of our work. Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

>>>>

\section{Discussion}\label{sec:discussion}

%Discuss the results. What is the outcome of your experimetns?
This subsection shortly discusses the empirical results and then extensively compares the findings of this work to the existing literature while trying to explain some of the differences. It is concluded by a subsection describing the limitations and possible extensions of this work.

\subsection{Our work}
This Section discusses our findings. First it will analyze the possible reasons behind the influence of input shuffling to the learning rate. Second it will conclude with the discussion of unexpected results.

\subsubsection{Shuffling and learning rate with SGD}
When the model used SGD as an optimizer it was able to use higher learning rates when input shuffling was used, as shown in figures \ref{fig:results_sgd}, \ref{fig:results_mom} and \ref{fig:results_nag}. Therefore arises the questions why these phenomena happen, especially as it did not happen with advanced optimizers. We attribute this phenomena to the use of the batched version of the SGNS.  In consequence to this batched version, when input shuffling is not used a lot of words will appear multiple times in the same batch, as explained in \ref{ssec:shuffling}. Therefore, the gradients will be an average of all the training samples. When the input is shuffled, less words will appear multiple time which will make the gradients more precise.  This is counter-attacked by advanced optimizers as they have adaptive learning rates, i.e a frequent appearing feature will have a low learning rate. Since high frequency words will appear more often multiple times in a batch, and therefore their average will not influence the results as much. The two above arguments could explain why a higher learning rate combined with SGD achieved better results with input shuffling but not when advanced optimizers are used.

@@

\subsection{Comparison to Gensim}
This Section will compare our finding extensively to Gensim \cite{gensim}. As explained in Section \ref{ssec:gensim}, Gensim is optimized to have a very high throughput, this made it possible to achieve a lot of computations. Furthermore, Gensim provides access to the loss and the resulting word embeddings, which facilitated the comparison process.
This Section is Structured as follows:First it will describe the used configuration of Gensim, second it will compare Gensim to our implementation of the SGNS with the use of SGD and finally compare our model with Adam to Gensim. A graphic showing the comparison of our models to Gensim can be found in Figure \ref{fig:gensim_vs_adam}.

@@

\subsubsection{Configuration of Gensim}
To compare our-self in a correct  manner we used the same dataset, with the same preprocessing parameters, i.e subsampling and min count. The hyper parameters of the network (window size, embedding size, number of negative sample, exponent to which the unigram distribution is raised to chose neg samples) are equivalent to ours, described in Section \ref{ssec:config}. The only difference with our model is the learning rate. Gensim has a starting learning rate of $0.025$ and linearly decrease it at every epoch to $0.0001$.

@@

\subsection{Future Work}
This work shows that the convergence time of the SGNS could be improved by using input shuffling and advanced optimizers. As with every work, there still exists possible extensions. First and foremost an aspect of our implementation that can be prejudicial is that we only extensively tested our model with one small dataset. The fact that we only used one dataset as well that it's a small one is problematic. \\ 
~~\textit{Problem with a small dataset:} \\ By using a very small dataset we do not use the model in the condition it is most needed for, as the dataset used in practice usually consists of more than 1 billion words. There is a small argument that can be made for machine translation as the use of small parallel corpus is not unusual in this field. \\
~~\textit{Problem with using only one dataset:}\\
It has been shown that some optimizers perform better with specific shapes of loss functions. To make a compelling argument it's, necessary to show that our model with the use of Input shuffling and Adam as its optimizer also outperform Gensim with other data sets, so that the claim can hold consistently.\\
Finally, our implementation did not outperform Gensim in run time, as this was not the goal of our work. Therefore, one could improve an already existing, optimized version, with input shuffling and advanced optimizers and should achieve a better run time than Gensim.

### 2019-05-07 icdm/submission/includes/evaluation.tex
discussion improved

\subsection{Configuration of the network}
The skip gram model has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter. The explanation of the parameters will be structured as follows:
\texttt{Parameter} - Description and tuning - \textit{Value}
\begin{itemize}
\item \texttt{Negative Samples:} Here we have to find a trade-off between, a parameter that is too high (better accuracy) and low ( faster computation time). For smaller data sets a higher number of negative samples is often needed. In their original paper Mikolov et al. \cite{mikolov2} recommend a value of 5-20. We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value. - \textit{10}
\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased. Mikolov et al. \cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5. - \textit{5}
\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}. We, therefore, used Gensim to find the best embedding possible. - \textit{100}
\item \texttt{Batch size}: As described in Section \ref{sec:contribution}, there is a trade off to find between quality and training time. We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
\end{itemize}

@@

\iffalse
\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
Embedding Size & Word Similarity on Gensim \\ \hline
50 & 0.65 \\ \hline
100 & 0.67 \\ \hline
200 & 0.65 \\ \hline
300 & 0.63 \\ \hline
\end{tabular}
\caption{Word similarity in relation to the size of the embedding}
\label{table:gensim_emb_size}
\end{table}
\fi
\subsection{Input Shuffling}
We used input shuffling as a technique to optimize the SGM.  We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGM. \\
Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in other to find the best optimum. Therefore, it's easier for the neural network to escape a local minimum. As for example if a network had converged to a local minimum after one epoch it probably cannot escape it. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the skip gram model. Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{sec:contribution}. When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples. But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller. In consequence the accuracy of the model should be optimized leading to a better convergence time. A definition of convergence time is given in the next section.

>>>>

\subsection{Configuration of the network}\label{ssec:config}
The skip gram model has a lot of possible parameters, that can be tuned. We experimented with different models and finally decided for one that we tried to optimize. This subsection will give a short overview of each parameter, where we will explain the process in which we chose the value of the given parameter. The explanation of the parameters will be structured as follows:
\texttt{Parameter} - Description and tuning - \textit{Value}
\begin{itemize}
\item \texttt{Negative Samples:} Here we have to find a trade-off between, a parameter that is too high (better accuracy) and low ( faster computation time). For smaller data sets a higher number of negative samples is often needed. In their original paper Mikolov et al. \cite{mikolov2} recommend a value of 5-20. We tested a few values in the range of 5 to 15, as 10 yielded state of the art results we chose this value. - \textit{10}
\item \texttt{Context Window:} The bigger the window the more training examples the network will have, but if the window is too big the semantic meaning of the window will be erased. Mikolov et al. \cite{mikolov} proposed a setting between 2-10, as all our sentences are of size 20 we chose 5. - \textit{5}
\item\texttt{ Dimension of the embedding}: Here the choice is less obvious, as the dimension needs to be high enough to capture the meaning, but cannot be too high as this leads to a decrease in performance as shown by Yin and Shen \cite{dimension_size}. We, therefore, used Gensim to find the best embedding possible. - \textit{100}
\item \texttt{Batch size}: As described in Section \ref{sec:contribution}, there is a trade off to find between quality and training time. We first used a batch size of 5000, but then decide after non conclusive results that 2000 would be better - \textit{2000}
\item \texttt{Alpha}: learning rate, this hyperparameter was tuned in every optimizer therefore only the range will be indicated - \textit{(1e-5,1)}
\end{itemize}

@@

\iffalse
\begin{table}[h]
\centering
\begin{tabular}{|l|l|}
\hline
Embedding Size & Word Similarity on Gensim \\ \hline
50 & 0.65 \\ \hline
100 & 0.67 \\ \hline
200 & 0.65 \\ \hline
300 & 0.63 \\ \hline
\end{tabular}
\caption{Word similarity in relation to the size of the embedding}
\label{table:gensim_emb_size}
\end{table}
\fi
\subsection{Input Shuffling}\label{ssec:shuffling}
We used input shuffling as a technique to optimize the SGM.  We will first describe input shuffling in a general way and then explain why we supposed that input shuffling could work well on the SGM. \\
Let $X = {x_1...x_n}$ be our input data set. Input Shuffling describes the process of taking a random permutation of the dataset as an input at each epoch.
The idea behind this technique is to present our optimizer with different  surfaces of our loss function, in other to find the best optimum. Therefore, it's easier for the neural network to escape a local minimum. As for example if a network had converged to a local minimum after one epoch it probably cannot escape it. But if we change the shape of the loss function, by input shuffling, then there would be a greater probability for the network to escape the local minimum.
\\
There are two reasons why we think that input shuffling is particularly well suited for the skip gram model. Firstly, we read our words sequentially, therefore,  words that only appear very early in the dataset will not benefit from the context words being already updated from others. The second idea is that we used the special batch technique described in Section \ref{sec:contribution}. When we do not use input shuffling the same word will appear with all of its context words as training samples in a batch, therefore the updae will be an average of all of these training samples. But if we would use input shuffling instead, then the probability of a word appearing multiple time in one batch would be smaller. In consequence the accuracy of the model should be optimized leading to a better convergence time. A definition of convergence time is given in the next section.
